\thispagestyle{fancy}

\textrm{}\\\\
\noindent\textbf{\huge\textsf{Acknowledgements}}\\\\

\noindent Over the past five years I have relied upon the friendship, guidance and support of many people to get me to this point and I would like to extend my gratitude toward them here.

First, a huge thank you to my supervisor, Dr. Christine Wilson.  I have had an amazing experience working with you and I am grateful for everything you have given me.  For your encouragement, ideas, knowledge and support.  For the many conversations we shared, both scientific and otherwise.  For giving me the opportunity to work with observations from Herschel, my dream dataset.  And for sending me to numerous places to share this work with others, and to the top of Mauna Kea!  No matter what lies ahead for me, I will always treasure my time here at McMaster, and all of the things I've learned along the way.

I would also like thank the numerous collaborators that have also helped me along the way.  To Dr. Suzanne Madden, my ``European Mom".  It's been six years since I first met you as a Master's student and now I've finally made it to the finish line!  Thank you for introducing me to Herschel and for keeping tabs on me after all these years.  Your friendship and support mean so much to me.

Thank you to Dr. Diane Cormier and Dr. Vianney Lebouteiller for showing me the ropes on PACS data processing and PACSman, to Dr. Kelly Foyle for your scientific help and advice, and to the rest of the Herschel SAG2 for helping me do the best science possible.  And many thanks to my supervisory committee members, Dr. Laura Parker and Dr. James Wadsley, for keeping me on track and for your suggestions and guidance on improving my research.

Tina, Cheryl, Rose and Mara.  You guys are the greatest secretaries a graduate student could ask for!  You are awesome!  I will always be thankful for the many conversations we've shared when I needed a break, as well as all of things you've done for me over the years.

I will be forever grateful for the friends I've made along the way.  To Rachel Ward-Maxwell, Aaron Maxwell, David Kirsh, Maximilien Schirm, Elizabeth Tasker, Damien Robertson, Jonathan Newton, Kevin Sooley and Leo Van Nierop.  You are all such amazing people and I can think of no other people I would have wanted to share the highs and lows of graduate school with.  I would be lost without your kindness, jokes, advice, support and friendship.  I will definitely miss our many crazy conversations over lunch!

I especially want to thank my parents for being there with me along this journey.  You have always been there for me, no matter what, and for that I will be eternally grateful.  I love you both so much!  And Lee, you are the best brother a sister could ask for!  You always put up with my craziness and lend an ear when I need one.

Lastly, to Kazimierz Sliwa.  Kaz, the past three years have been amazing and I am so thankful that I met you.  You have always believed in me, even when I didn't believe in myself, and I could not have made it without your love and support.  And now it's my turn to be there for you while you finish your own degree.  I love you!


