\thispagestyle{fancy}

\textrm{}\\\\
\noindent\textbf{\huge\textsf{Co-Authorship}}\\\\

\noindent 
Chapters~\ref{chapter2}, \ref{chapter3}, and \ref{chapter4} of this thesis contain original scientific research written by myself, Tara Jill Parkin.  Chapter~\ref{chapter2} has been published as a peer-reviewed journal article in the Monthly Notices of the Royal Astronomical Society (MNRAS).  The reference to this work is:

\noindent \emph{Parkin, T.~J., Wilson, C.~D., Foyle, K., Baes, M., Bendo, G.~J., Boselli, A., Boquien, M., Cooray, A., Cormier, D., Davies, J.~I., Eales, S.~A., Galametz, M., Gomez, H.~L., Lebouteiller, V., Madden, S., Mentuch, E., Page, M.~J., Pohlen, M., Remy, A., Roussel, H., Sauvage, M., Smith, M.~W.~L., Spinoglio, L., 2012, MNRAS, Volume 422, Issue 3, pp.~2291--2301.}

\noindent My supervisor, Dr. Christine Wilson, is the second author.  Dr. Kelly Foyle ran my photometry through her dust spectral energy distribution code for the purposes of evaluating the impact of having two free parameters versus one on my results; however, we used my model for the full analysis.  Dr. Foyle also provided general comments on the paper.  Dr. Marc Sauvage, Dr. Helen Roussel, Dr. Michael Pohlen, and Matthew W.~L. Smith processed the \emph{Herschel} PACS and SPIRE photometric maps for Cen~A, and Dr. George Bendo provided the convolution kernels for the PACS photometry.  It is the policy of the SPIRE Special Astronomy Group 2 (SAG2) collaboration, of which I am a member, to circulate drafts of articles to be submitted for publication to other members for an internal peer-review process.  Any member that provides comments on a circulated draft is automatically added to the list of authors on that paper.  Thus, the remaining co-authors of this paper are members of the SAG2 that provided me with comments and suggestions to improve the draft prior to submission, but contributed nothing further to the primary analysis.  

At the time of my final thesis submission to the School of Graduate Studies, the work presented in Chapter~\ref{chapter3} has been accepted for publication by the peer-reviewed Astrophysical Journal (ApJ); however, it has not yet been published and thus does not yet have a full bibliographic reference.  The author list is as follows:

\noindent \emph{T.~J. Parkin, C.~D. Wilson, M.~R.~P. Schirm, M. Baes, M. Boquien, A. Boselli, A. Cooray, D. Cormier, K. Foyle, O.~\L. Karczewski, V. Lebouteiller, I. de~Looze, S.~C. Madden, H. Roussel, M. Sauvage and L. Spinoglio.}

\noindent Again, Dr. Christine Wilson is second author.  Maximilien Schirm processed the SPIRE FTS spectroscopic observations and produced the [N~\textsc{ii}](205~$\mu$m) map of M51 for use in my analysis.  Dr. Marc Sauvage and Dr. Helen Roussel processed the PACS photometry I used as part of my analysis.  Dr. Diane Cormier and Dr. Vianney Lebouteiller provided me with the code PACSman, which I used to process the PACS spectroscopy.  The remaining co-authors are members of SAG2 who provided comments upon circulating a draft of the paper and obtained co-authorship as described above.

The research presented in Chapter~\ref{chapter4} is in preparation to be submitted to ApJ. It has been circulated to the SAG2 for comments but they have not been incorporated into the work as of the time of final submission.  The author list is as follows:

\noindent \emph{T.~J. Parkin, C.~D. Wilson, M.~R.~P. Schirm, M. Baes, M. Boquien, A. Boselli, D. Cormier, M. Galametz, O.~\L. Karczewski, V. Lebouteiller, S. Madden, H. Roussel, M.~W.~L. Smith, and L. Spinoglio}

\noindent Dr. Christine Wilson is my supervisor and is second author.  Maximilien Schirm processed the SPIRE FTS spectroscopy and produced the [N~\textsc{ii}](205~$\mu$m) map for Cen~A.  Dr. Helen Roussel processed the PACS photometry I used for my analysis.  Dr. Diane Cormier and Dr. Vianney Lebouteiller provided me with PACSman, and the remaining co-authors are members of SAG2 who provided comments upon circulating a draft of the paper and obtained co-authorship as described above.

All previously published material has been reformatted to conform to the required thesis 
style.  I grant an irrevocable, non-exclusive license to McMaster University and the National Library of Canada to reproduce this material as part of this thesis.
\newpage
\thispagestyle{empty}
\mbox{}